---
title: "Network Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{network tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(enet)
```

## Network Basics

### Vertex and edge

Each node is a vertex and the connection between nodes is a edge in the network. The connection can be directed or undirected depending on the relationship.

### Build the network

Adjacency matrices were always used to build the network. It's a square matrix with n dimensions. Row i and column j is equal to 1 if and only if vertices i and j are connected. In directed network, such values could be 1 for i to j and -1 for j to i. 

### Network attributes

Vertex/edge attributes could be the group information or metadata about the nodes/connections. The edges could be weighted as attribute.

Path is the way from one node to another node in the network and you could find the shortest path in the path. The largest distance of a graph is called its diameter.

An undirected network is connected if there is a way from any vertex to any other. Connected networks can further classified according to the strength of their connectedness. An undirected network with at least two paths between each pairs of nodes is said to be biconnected. 

The transitivity of network is a crude summary of the structure. A high value means that nodes are connected well locally with dense subgraphs. Network data sets typically show high transitivity.

Maximum flows and minimum cuts could be used to check the largest volumns and smallest path flow between two nodes. For example, two hubs is connected by one node and the largest volumn and smallest path flow between two nodes from each hub could be counted at the select node.

Sparse network has similar number of edges and the number of nodes. Dense network has the number of edges as a quadratic function of the nodes.
